---
title: "Covid91 vaccine study Final2021"
author: "Logan DeHay"
date: "November 23 2021"
output:
  pdf_document: default
  toc: yes
  html_document: null
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))

FinalData <- FinalData %>% mutate_if(is.character, as.factor)
FinalData$treatment <- relevel(FinalData$treatment, ref = "placebo")
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Guys

```{r}
Guys <- filter(FinalData, sex=="guy")
```
# Graphical Results
```{r}
barchartGC(~infected + treatment,data=Guys)
barchartGC(~infected + treatment,data=Guys, type="percent")

```
  
The data illustrates that the overwhelming majority of the population of Guys were not infected, however the guys who received the treatment, actually were more likely to acquire COVID than the ones who received the placebo.
  
# Numerical Results
```{r}

table1 <- xtabs(~infected + treatment, data=Guys)
rowPerc(table1)
colPerc(table1)
```
  
The numerical summary states that the vast majority of the guys were ok and uninfected. Of those that were, 58.16% had received the drug while 41.84% had received the placebo. 
  
# Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Guymod <- fisher.test(table1)
guyeff <- (Guymod$estimate/(Guymod$estimate+1))*100
Guyeffupper <- Guymod$conf.int[2]
Guyeffuppereff <- (Guymod$conf.int[2]/(Guymod$conf.int[2]+1))*100
Guyefflower <- Guymod$conf.int[1]
Guyefflowereff <- (Guymod$conf.int[1]/(Guymod$conf.int[1]+1))*100
options(digits=3)
```
  

  The chi square test shown above demonstrates a sizable difference in the expected counts by the null and the actual results in the data. Not only this, it has a small p-value below 0.05 which is the probability under our null hypothesis conditions that there is no significant difference between receiving the placebo and the treatment. The Fisher test similarly had a small p-value like the chi square test and as a result, we can assert that the drug taken has a significant effect on the likelihood of guys becoming infected with COVID or not. The Fisher test calculated that the odds of acquiring COVID when taking the drug is `r Guymod$estimate` times more likely than if not. Our odds ratio can be seen above at 1.473:1. This is based in a 95% confidence interval ranging from 1.1767 to 1.8497. We can reject our null hypothesis according to the tests above. The efficacy of the drug for the guys is `r guyeff`% with a 95% confidence interval of `r Guyefflowereff` to `r Guyeffuppereff`. As is evident by these values, the chance of men getting COVID increases when taking the drug, and as such, I am not recommending this treatment for guys.

  
# Gals

```{r}
Gals <- filter(FinalData, sex=="gal")
```

# Graphical Results
```{r}
barchartGC(~infected + treatment,data=Gals)
barchartGC(~infected + treatment,data=Gals, type="percent")
```

Our frequency bar chart displays the normality that the large majority of the sample were OK, while a minority contracted COVID. The percent bar chart illustrates that comparatively, there was a discernible difference between the treatment and the placebo. About 30% of the gal population that contracted COVID-19 received the drug while the other 70% of that population received the placebo. This suggests that the study drug Covid91 could be effective for gals because it seemingly decreased the likelihood of contarcting COVID-19. 

# Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=Gals)
rowPerc(table1)
colPerc(table1)
```
  
The first table shows that over 90% of the population were totally fine disregarding drug or placebo. Of those that became infected, 66.9% had received the placebo while only 33.1% of the infected were those that received the drug. Akin to our graphical results, the discrepancy between the two groups indicates a bias in the data.
  
# Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Galmod <- fisher.test(table1)
galeff <- (Galmod$estimate/(Galmod$estimate+1))*100
Galeffupper <- Galmod$conf.int[2]
Galeffuppereff <- (Galmod$conf.int[2]/(Galmod$conf.int[2]+1))*100
Galefflower <- Galmod$conf.int[1]
Galefflowereff <- (Galmod$conf.int[1]/(Galmod$conf.int[1]+1))*100
options(digits=3)
```
  
Our alternative hypothesis is that there is a tangible difference in the probability of contracting COVID-19 dependent on whether a person had taken or not taken the study drug treatment. The null hypothesis is that there is no difference between those that received the  drug Covid91 and those that received a placebo in getting infected with COVID. The Chi-squared  and Fisher tests similarly calculated p-values lower than 0.05, indicating that we can determine the difference significant in our data. The fisher p-value was 2e-16 and the Chi Square p-value 4e-16. The probability of our results being unbiased and naturally occurring is extremely unlikely. The Fisher test calculated that the odds ratio of women contracting COVID-19 when taking the treatment was `r Galmod$estimate` times less than those who received the placebo. As such, we can reject the null hypothesis because women are less likely to get infected with COVID after getting the treatment. The efficacy of the drug for the gals is `r galeff`% with a 95% confidence interval of `r Galefflowereff` to `r Galeffuppereff`. Based on this, I would advise gals to get the treatment.
  
# LGBTQ

```{r}
LGBTQ <- filter(FinalData, LGBTQ=="gay")
```

# Graphical Results
```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")
```
  
The frequency bar chart compares the frequency of the LGBTQ population who contracted COVID-19 and are ok. A overwhelmming majority of the population are ok with a small portion becoming infected. Our percent bar chart shows that of those who were infected, around 75% had received the treatment with the other 25% getting the placebo. Something is fishy! Our graphical results indicate a bias to the data.
   
# Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```
  
  The table at the bottom represents the percentage of the LGBTQ population who were ok and infected total. So 92.03% of those who received the drug were fine while 96.64% of those who received the placebo were OK. The top table represents what percentage of the infected population had taken the rug or the placebo. Our "OK" population was equally composed of those who had taken and did not take the drug. However, of the infected population, 70.7% had received the drug (a large correlation). This fits our graphical results.
   
# Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
LGBTQmod <- fisher.test(table1)
LGBTQeff <- (LGBTQmod$estimate/(LGBTQmod$estimate+1))*100
LGBTQeffupper <- LGBTQmod$conf.int[2]
LGBTQeffuppereff <- (LGBTQmod$conf.int[2]/(LGBTQmod$conf.int[2]+1))*100
LGBTQefflower <- LGBTQmod$conf.int[1]
LGBTQefflowereff <- (LGBTQmod$conf.int[1]/(LGBTQmod$conf.int[1]+1))*100
options(digits=3)
```
  
Our Chi-Squared test displayed a large variation between the expected counts by the null hypothesis and our actual counts. The varying values can be seen above. Our null hypothesis is that the treatment has no significant effect on the likelihood of becoming infected with COVID-19. Our p-values are significantly low for both the Chi-squared test and the fisher exact test (1e-05 and 6e-06). This is lower than our probability under the null hypothesis of 0.05. The Fisher test calculated the odds ratio of LGBTQ people getting infected with COVID when taking the treatment to be 2.49 to 1. Based on our data, LGBTQ people are `r LGBTQmod$estimate` times more likely to become infected if they receive treatment. As is evident by our calculations, the null hypothesis can be reject, and it can be asserted that the drug COVID91 has a significant negative effect on the probability of getting COVID-19. The efficacy of the drug when taken by LGBTQ people is `r LGBTQeff`%, with a 95% confidence interval of `r LGBTQefflowereff` to `r LGBTQeffuppereff`. The drug would have a negative impact on this population, and based on our p0values, efficacy, and odds ratio, I would not recommend this drug to be taken by LGBTQ citizens.
  
   
# Drug Users
  
```{r}
DrugUsers <- filter(FinalData, DrugUser=="yes")
```

# Graphical Results
```{r}
barchartGC(~infected + treatment,data=DrugUsers)
barchartGC(~infected + treatment,data=DrugUsers, type="percent")
```
  
The frequency bar chart above shows that out of thousands of people who received the treatment and placebo, less than 200 total became infected. However, there seems to be more infected people when they received the placebo as opposed to the drug. In our percent bar chart, around 35% of the drug users who became infected had been given the drug, while a whopping 70% had been given the placebo. This indicates a statistical bias which can be possibly derived from having received the treatment.
  
# Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=DrugUsers)
rowPerc(table1)
colPerc(table1)
```
  
The bottom portion of our tables shows that 3.63% of the population who had been given the drug got COVID-19 while 6.46% of the population who had been given the placebo got COVID. Of those who had become infected, 66.7% of them were given the placebo. Similarly to our graphical results, the numerical results indicate a correlation between contraction of the virus, and ingestion of the drug.
  
# Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
DrugUsersmod <- fisher.test(table1)
DrugUserseff <- (DrugUsersmod$estimate/(DrugUsersmod$estimate+1))*100
DrugUserseffupper <- DrugUsersmod$conf.int[2]
DrugUserseffuppereff <- (DrugUsersmod$conf.int[2]/(DrugUsersmod$conf.int[2]+1))*100
DrugUsersefflower <- DrugUsersmod$conf.int[1]
DrugUsersefflowereff <- (DrugUsersmod$conf.int[1]/(DrugUsersmod$conf.int[1]+1))*100
options(digits=3)
```
  
The Peason Chi Squared test at the top has a p-value of 0.003 which is relatively smaller than 0.05. Our counts expected by the null and observed counts vary slightly form each other, but they vary less exaggeratedly than the LGBTQ counts. Our Fisher exact test has a p-value of 0.002, which is also relatively smaller than 0.05. As a result, we can effectively claim the null hypothesis to be false, an that the intake of our drug has a significant effect on the chances of drug users becoming ill with COVID-19. Our odds ratio was 0.546:1, meaning that Drug Users were `r DrugUsersmod$estimate` times less likely to get COVID-19 because they received the study drug. Our tests and results prove that Drug Users are less likely to become COVID positive if they receive the treatment. The efficacy of the drug for the Drug Users is `r DrugUserseff`% with a 95% confidence interval of `r DrugUsersefflowereff` to `r DrugUserseffuppereff`. Taking all of our data and results into account, I would recommend this droog to any Drug Users because it has a positive effect on their likelihood of not contracting COVID.
  
# Discussion/Conclusion

