---
title: "Covid91 vaccine study Final2021"
author: "Logan DeHay"
date: "November 23 2021"
output:
  pdf_document: default
  toc: yes
  html_document: null
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))

FinalData <- FinalData %>% mutate_if(is.character, as.factor)
FinalData$treatment <- relevel(FinalData$treatment, ref = "placebo")
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Guys

```{r}
Guys <- filter(FinalData, sex=="guy")
```
# Graphical Results
```{r}
barchartGC(~infected + treatment,data=Guys)
barchartGC(~infected + treatment,data=Guys, type="percent")

```
  
The data illustrates that the overwhelming majority of the population of Guys were not infected, however the guys who received the treatment, actually were more likely to acquire COVID than the ones who received the placebo.
  
# Numerical Results
```{r}

table1 <- xtabs(~infected + treatment, data=Guys)
rowPerc(table1)
colPerc(table1)
```
  
The numerical summary states that the vast majority of the guys were ok and uninfected. Of those that were, 58.16% had received the drug while 41.84% had received the placebo. 
  
# Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Guymod <- fisher.test(table1)
guyeff <- (Guymod$estimate/(Guymod$estimate+1))*100
Guyeffupper <- Guymod$conf.int[2]
Guyeffuppereff <- (Guymod$conf.int[2]/(Guymod$conf.int[2]+1))*100
Guyefflower <- Guymod$conf.int[1]
Guyefflowereff <- (Guymod$conf.int[1]/(Guymod$conf.int[1]+1))*100
options(digits=3)
```
  

  The chi square test shown above demonstrates a sizable difference in the expected counts by the null and the actual results in the data. Not only this, it has a small p-value below 0.05 which is the probability under our null hypothesis conditions that there is no significant difference between receiving the placebo and the treatment. The Fisher test similarly had a small p-value like the chi square test and as a result, we can assert that the drug taken has a significant effect on the likelihood of guys becoming infected with COVID or not. The Fisher test calculated that the odds of acquiring COVID when taking the drug is `r Guymod$estimate` times more likely than if not. This is based in a 95% confidence interval ranging from 1.1767 to 1.8497. We can reject our null hypothesis according to the tests above. The efficacy of the drug for the guys is `r guyeff`% with a 95% confidence interval of `r Guyefflowereff` to `r Guyeffuppereff`. As is evident by these values, the chance of men getting COVID increases when taking the drug, and as such, I am not recommending this treatment for guys.

  
# Gals

```{r}
Gals <- filter(FinalData, sex=="gal")
```

# Graphical Results
```{r}
barchartGC(~infected + treatment,data=Gals)
barchartGC(~infected + treatment,data=Gals, type="percent")
```

Our frequency bar chart displays the normality that the large majority of the sample were OK, while a minority contracted COVID. The percent bar chart illustrates that comparatively, there was a discernible difference between the treatment and the placebo. About 30% of the gal population that contracted COVID-19 received the drug while the other 70% of that population received the placebo. This suggests that the study drug Covid91 could be effective for gals because it seemingly decreased the likelihood of contarcting COVID-19. 

# Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=Gals)
rowPerc(table1)
colPerc(table1)
```
  
The first table shows that over 90% of the population were totally fine disregarding drug or placebo. Of those that became infected, 66.9% had received the placebo while only 33.1% of the infected were those that received the drug. Akin to our graphical results, the discrepancy between the two groups indicates a bias in the data.
  
# Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Galmod <- fisher.test(table1)
galeff <- (Galmod$estimate/(Galmod$estimate+1))*100
Galeffupper <- Galmod$conf.int[2]
Galeffuppereff <- (Galmod$conf.int[2]/(Galmod$conf.int[2]+1))*100
Galefflower <- Galmod$conf.int[1]
Galefflowereff <- (Galmod$conf.int[1]/(Galmod$conf.int[1]+1))*100
options(digits=3)
```
  
Our alternative hypothesis is that there is a tangible difference in the probability of contracting COVID-19 dependent on whether a person had taken or not taken the study drug treatment. The null hypothesis is that there is no difference between those that received the  drug Covid91 and those that received a placebo in getting infected with COVID. The Chi-squared  and Fisher tests similarly calculated p-values lower than 0.05, indicating that we can determine the difference significant in our data. The fisher p-value was 2e-16 and the Chi Square p-value 4e-16. The probability of our results being unbiased and naturally occurring is extremely unlikely. The Fisher test calculated that the odds ratio of women contracting COVID-19 when taking the treatment was `r Galmod$estimate` times less than those who received the placebo. As such, we can reject the null hypothesis because women are less likely to get infected with COVID after getting the treatment. The efficacy of the drug for the gals is `r galeff`% with a 95% confidence interval of `r Galefflowereff` to `r Galeffuppereff`. Based on this, I would advise gals to get the treatment.
  
# LGBTQ

# Drug Users

# Discussion/Conclusion

